Hydra Battles and Cie

### 2021-01-28 doc/chapter-primrec.tex
add an exercise
add a new exercise
additions to an exercise on Ackermann function

\begin{Coqsrc}
Fixpoint fac n :=
  match n with 
          | 0 => 1
          | S p  => n * fac p
  end.

@@

\textbf{Hint:} You may have to look again at the lemmas of the library
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec} if you meet some difficulty.
\end{exercise}

@@

@@

\index{Exercises}
\begin{exercise}
Show that the function \texttt{min: naryFunc\,2} is primitive recursive.
\end{exercise}

@@

\begin{enumerate}
\item Show that, for any $n$, the function $A\,n$ is primitive recursive.
\item Show that for all $n$ and $m$, $A\,n\,m > m$ and  that the functions $A\,n$ are strictly monotonous.
Show also that, for any $n$ and $m$, $A\,n\,m \leq A\,(n+1)\,m$.
         Finally, show that for any $n$, $m$ and $k$, 
$A\,k\,(A\,m\,n)\leq A\,(2+\textrm{max}(k,m))\,n$.
\item Show that for any primitive recursive function $f$ of $k$ arguments,
for any $x_1,\dots\,x_k$, $f(x_1,\dots,x_k)\leq A\,n\,(x_1+\dots+x_k)$.
\item Show that the Ackermann function is not primitive recursive.

@@

\begin{remark}
Note that the Ackermann function is a counter-example to the (false) statement:
\begin{quote}
{\color{red}
  ``Let $f$ be a function of type \texttt{naryFunc\,2}. If, for any $n$, the fonction $f(n)$ is primitive recursive, then f is primitive recursive.''}
\end{quote}
\end{remark}
\end{project}

>>>>

\begin{Coqsrc}
Fixpoint fact n :=
  match n with 
          | 0 => 1
          | S p  => n * fact p
  end.

@@

\textbf{Hint:} You may have to look again at the lemmas of the library
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec} if you meet some difficulty.

\emph{You may start this exercise with the file
\url{../exercises/primrec/MorePRExamples.v}.}
\end{exercise}

@@

\emph{You may start this exercise with
\url{../exercises/primrec/FibonacciPR.v}.}

@@

\index{Exercises}
\begin{exercise}
Show that the function \texttt{min: naryFunc\,2} is primitive
recursive.

\emph{You may start this exercise with
\url{../exercises/primrec/MinPR.v}.}

@@

\begin{enumerate}
\item Show that, for any $n$, the function $A\,n$ is primitive recursive.
\emph{\textbf{Done.}}

\item Show that for all $n$ and $m$, $A\,n\,m > m$ and  that the functions $A\,n$ are strictly monotonous.
\emph{\textbf{To do.}}

Show also that, for any $n$ and $m$, $A\,n\,m \leq A\,(n+1)\,m$. \emph{\textbf{To do.}}
         Finally, show that for any $n$, $m$ and $k$, 
$A\,k\,(A\,m\,n)\leq A\,(2+\textrm{max}(k,m))\,n$.
\emph{\textbf{To do.}}
\item Show that for any primitive recursive function $f$ of $k$ arguments,
for any $x_1,\dots\,x_k$, $f(x_1,\dots,x_k)\leq A\,n\,(x_1+\dots+x_k)$. \emph{\textbf{To do.}}
\item Show that the Ackermann function is not primitive recursive. \emph{\textbf{To do.}}

@@

\emph{You may start this exercise with
\url{../exercises/primrec/AckermannNotPR.v}.}

\begin{remark}
Note that the Ackermann function is a counter-example to the (false) statement:
\begin{quote}
{\color{red}
  ``Let $f$ be a function of type \texttt{naryFunc\,2}. If, for any $n$, the fonction $f(n)$ is primitive recursive, then f is primitive recursive.''}
\end{quote}
\end{remark}

\end{project}
